---
title: "Dataviz"
author: "Daniel Kreuz"
date: "9 11 2022"
output:
  html_document: default
  pdf_document: default
---

```{r Imports}

source("../functions/helpers.R")

library(data.table)
library(tidyverse)
library(magrittr)
library(checkmate)
library(plotly)
library(clinUtils)
library(janitor)
library(rcompanion)
library(DescTools)
library(FSelectorRcpp)

```

```{r data}

train <- read.csv("../data/training_set_values.csv", stringsAsFactors = TRUE) %>% setDT()
train_labels <- read.csv("../data/training_set_labels.csv", stringsAsFactors = TRUE) %>% setDT()
test <- read.csv("../data/test_set_values.csv", stringsAsFactors = TRUE) %>% setDT()
submission <- read.csv("../data/SubmissionFormat.csv")

```

```{r First impression}

#Check the data types
glimpse(train)

##### Results
# date_recorded should be of type date
# num_private ??
# public_meeting should be of type bool
# permit should be of type bool
# district code should be of type factor
# region code should be of type factor

# Change Datatypes via function

train <- change_datatypes(data = train)

assert_logical(train[,public_meeting])
assert_logical(train[,permit])
assert_date(train[,date_recorded])

```

```{r Target distribution}

tgt_dist <- plot_ly(data = train_labels, x = ~status_group) %>%
  layout(title = "Absolute distribution of the trainings target variable",
         yaxis = list(title = "Absolute"))

relative_train_labels <- train_labels[, .N, by = .(status_group)]
relative_train_labels[, rel := round(N / sum(N), digits = 2)]

rel_tgt_dist <- plot_ly(data = relative_train_labels, x = ~status_group, y = ~rel) %>% 
  layout(title = "Relative distribution of the trainings target variable",
         yaxis = list(title = "Relative"))

tgt_dist
rel_tgt_dist



```

We have a pretty unbalanced distribution in our target variable.
While the two classes functional and non functional are mostly evenly distributed we have only very few cases (~ 7%) of functional pumps that need repair. 
This could later lead to problems if our test set has more functional pumps that need repair and our model is not adjusted to this case. 




```{r Numerical Analysis}

# Cols to review in the numerical analysis
# amount_tsh, gps_height longitude, latitude, num_private, population

plot_ly(data = train, x = ~log(amount_tsh), type = "histogram") %>%
  layout(title = "Log Amount of flow in pumps")
plot_ly(data = train, x = ~gps_height, type = "histogram") %>%
  layout(title = "The Height of the pumps")
plot_ly(data = train, x = ~longitude, type = "histogram") %>%
  layout(title = "Longitude")
plot_ly(data = train, x = ~latitude, type = "histogram") %>%
  layout(title = "Latitude")
plot_ly(data = train, x = ~num_private, type = "histogram") %>%
  layout(title = "Num Private (Unknown)")
plot_ly(data = train, x = ~log(population), type = "histogram") %>%
  layout(title = "Log Population")


```

```{r Categorical Analysis, results='asis'}

# Plot a barchart for every factor value

fact_cols <- names(Filter(is.factor, train))
i <- 0
plots <- list()


for (col in fact_cols){
  p <- plot_ly(data = train, x = ~get(col), type = "histogram") %>%
    layout(xaxis = list(title = col))
  print(p)
}
  




```


## funder
We have a lot of different funders for all the pumps. The government is funding most of the pumps while other funders only have a few.
It may be possible that the pumps funded by the government are more functional than the pumps funded by private(?) companies
### Hypothesis 1: Pumps by the government are relatively more functional than the pumps funded by other contractors.

## Installer
Like the funders. The installer DWE has installed most of the pumps. 3655 pumps have no installer. We could have a look how the DWS installer influences the functionality.
Missing Data

## wpt_name
wpt_name is the name of the waterpoint if there is one. We can see that some waterpoints are more common than others. We should not focus on the insight here.

## basin
We only have a few basins which are mostly evenly distributed. We could try to get more insights later here. Maybe find association with other variables
### Hypothesis 2: The basin is associated with other variables in the dataset (Cramér's V)

## subvillage
We have a lot of different subvillage in our dataset. No insight to gain here.

## region
In comparison to the subvillage, we have fewer regions. This could come in handy, when we don´t want to use the subvillage in our data
### Hypothesis 3: The subvillage does not bring more information to the table than the region

## region_code
The region code is a little more accurate than the region. We should have a look if there is a difference in information gain later on.

## district_code
Similar to region code

## lga
Also geographic location. Similar to ther geographic informations in the dataset.

## ward
Also geographic location. Similar to ther geographic informations in the dataset.

## recorded_by
Only one value. Can be deleted

## scheme_management
Who operates the waterpoint. VWC has a lot more waterpoints than the rest of managements. We could have a look if there is a difference in functionality
### Hypothesis 4: VWC's management has an major influence on the functionality of waterpoints.

## scheme_name: 
A lot of different management companies. Too many to gain information. 
Can be deleted

## extraction_type
Different groups of extraction types. We could get information about the functionality by extraction type

## extraction_type_group
See extraction_type

## extraction_type_class
See extraction_type
### Hypothesis 5: The functionality is evenly distributed between all the extraction type columns 

## management
Like the scheme_management

## management_group
More generalized like that management

## payment
Information how the water is paid for.

## payment_type
Should be very similar to payment. 
### Hypothesis 6: The payment and payment time is redundand.

## water_quality
Mostly soft water in the wells. Sometimes salty. Rest is distributed

## quality_group
A little more generalized than the water_quality

## quantity
Information about the quantity of water. We should see if the quanitity is changing whether the pump is functional or not.
### Hypothesis 7: The quanity is associated to the functionality of the pump.

## quantity_group
Same as quantity. Can be deleted

## source
Where the water is coming from. Mostly from river, shallow well, spring or machines

## source_type
More generalized than source

## source_class
More generalized than source_type

## waterpoint_type
What type of waterpoint it is. Need to get information from the internet or stakeholders on what influence this could have

## waterpoint_type_group
Mostly the sameas waterpoint_type. A little more generalized


```{r Join labels and delete recorded_by}

train <- train[train_labels, on = "id"]
train <- train[,recorded_by := NULL]


```




# Hypotheses

## Hypothesis 1: Pumps by the government are relatively more functional than the pumps funded by other contractors.

```{r Hypo 1}

hypo_1 <- copy(train)
hypo_1[, .N, by = .(funder)][order(-N)]
hypo_1[funder != "Government Of Tanzania", funder := "Other"]
melted <- hypo_1[, .N, by = .(funder, status_group)][, rel := N / sum(N), by = .(funder)]
casted <- dcast(melted, funder ~ status_group, value.var = "rel")
casted <- casted %>%
  droplevels() %>%
  clean_names()


fig <- plot_ly(data = casted, x = ~funder, y = ~functional, type = "bar", name = "functional") %>%
  add_trace(y = ~functional_needs_repair, name = "functional needs repair") %>%
  add_trace(y = ~non_functional, name = "non functional") %>%
  layout(yaxis = list(title = "Relative count"), barmode = "stack")

fig
```

Apparently we have more functional pumps with other contractors. Lets adjust the graphic to see the top 20 contractors with their relative functionalities.

```{r Hypo 1 Addition}

hypo_1_add <- copy(train)
hypo_1_add[, .N, by = .(funder)][order(-N)]
top_funders <- hypo_1_add[, .N, by = .(funder)][order(-N)][1:20][,funder]
hypo_1_add[!funder %in% top_funders, funder := "Other"]
melted <- hypo_1_add[, .N, by = .(funder, status_group)][, rel := N / sum(N), by = .(funder)]
casted <- dcast(melted, funder ~ status_group, value.var = "rel")
casted <- casted %>%
  droplevels() %>%
  clean_names()


fig <- plot_ly(data = casted, x = ~funder, y = ~functional, type = "bar", name = "functional") %>%
  add_trace(y = ~functional_needs_repair, name = "functional needs repair") %>%
  add_trace(y = ~non_functional, name = "non functional") %>%
  layout(yaxis = list(title = "Relative count"), barmode = "stack")

fig
```

We can see only a slight difference in the contractors. Interesting is that the waterpoints funded by the german republic is mostly functional while the others are mostly under 50%.
We keep this in mind for later steps. We could categorize the different funders in our preparation.


## Hypothesis 2: The basin is associated with other variables in the dataset (Cramér's V)

```{r}

fact_cols <- names(Filter(is.factor, train))

for(col in fact_cols[fact_cols != "basin"]){
  cramer <- cramerV(x = train$basin, y = train[[col]])
  chisq <- chisq.test(x = train$basin, y = train[[col]])
  print(paste0("Cramer V for basin ~ ", col, " = ", cramer ," , p-Value = ", chisq["p.value"]))
}

```

The Cramer's V shows that we have a strong association between the basin and our variables lga und ward. We can therefore keep only one variable.
We could use some techniques later to find out which variable has more information for our target (Maybe PCA, TheilsU?)

## Hypothesis 3: The subvillage does not bring more information to the table than the region

```{r}

ig_subv <- information_gain(status_group ~ subvillage, data = train)
ig_region <- information_gain(status_group ~ region, data = train)
cv <- cramerV(x = train$subvillage, y = train$region)

print(paste0("Information Gain for Subvillage: ", ig_subv$importance, ". Information Gain for Region: ", ig_region$importance, ". CramersV: ", cv))



```

We have a much higher information gain for the subvillage in comparison to the region. BUT we also have way more distinct subvillages. Therefore the information gain is not really meaningful.
I additionally calculated the CramersV again to see if we have a strong association between those variables. As expected we have. Therefore we later need to see, if we can get rid of one of those variables.


### Hypothesis 4: VWC's management has an major influence on the functionality of waterpoints.

```{r}

hypo_4 <- copy(train)
hypo_4[, .N , by = .(scheme_management)][order(-N)]
hypo_4[scheme_management != "VWC", scheme_management := "Other"]
melted <- hypo_4[, .N, by = .(scheme_management, status_group)][, rel := N / sum(N), by = .(scheme_management)]
casted <- dcast(melted, scheme_management ~ status_group, value.var = "rel")
casted <- casted %>%
  droplevels() %>%
  clean_names()


fig <- plot_ly(data = casted, x = ~scheme_management, y = ~functional, type = "bar", name = "functional") %>%
  add_trace(y = ~functional_needs_repair, name = "functional needs repair") %>%
  add_trace(y = ~non_functional, name = "non functional") %>%
  layout(yaxis = list(title = "Relative count"), barmode = "stack")

fig





```


We have no visible difference between the management of VWC and others.


## Hypothesis 5: The functionality is evenly distributed between all the extraction type columns

```{r}

hypo_5 <- copy(train)
hypo_5[, .N , by = .(extraction_type)][order(-N)]
melted <- hypo_5[, .N, by = .(extraction_type, status_group)][, rel := N / sum(N), by = .(extraction_type)]
casted <- dcast(melted, extraction_type ~ status_group, value.var = "rel")
casted <- casted %>%
  droplevels() %>%
  clean_names()


fig_1 <- plot_ly(data = casted, x = ~extraction_type, y = ~functional, type = "bar", name = "functional") %>%
  add_trace(y = ~functional_needs_repair, name = "functional needs repair") %>%
  add_trace(y = ~non_functional, name = "non functional") %>%
  layout(yaxis = list(title = "Relative count"), barmode = "stack")

###

hypo_5 <- copy(train)
hypo_5[, .N , by = .(extraction_type_group)][order(-N)]
melted <- hypo_5[, .N, by = .(extraction_type_group, status_group)][, rel := N / sum(N), by = .(extraction_type_group)]
casted <- dcast(melted, extraction_type_group ~ status_group, value.var = "rel")
casted <- casted %>%
  droplevels() %>%
  clean_names()


fig_2 <- plot_ly(data = casted, x = ~extraction_type_group, y = ~functional, type = "bar", name = "functional") %>%
  add_trace(y = ~functional_needs_repair, name = "functional needs repair") %>%
  add_trace(y = ~non_functional, name = "non functional") %>%
  layout(yaxis = list(title = "Relative count"), barmode = "stack")


###


hypo_5 <- copy(train)
hypo_5[, .N , by = .(extraction_type_class)][order(-N)]
melted <- hypo_5[, .N, by = .(extraction_type_class, status_group)][, rel := N / sum(N), by = .(extraction_type_class)]
casted <- dcast(melted, extraction_type_class ~ status_group, value.var = "rel")
casted <- casted %>%
  droplevels() %>%
  clean_names()


fig_3 <- plot_ly(data = casted, x = ~extraction_type_class, y = ~functional, type = "bar", name = "functional") %>%
  add_trace(y = ~functional_needs_repair, name = "functional needs repair") %>%
  add_trace(y = ~non_functional, name = "non functional") %>%
  layout(yaxis = list(title = "Relative count"), barmode = "stack")


fig_1
fig_2
fig_3





```

We can see that with the extraction type "other" has mostly non-functional waterpoints. Other than that the distribution is kind of even.


## Hypothesis 6: The payment and payment time is redundand.

```{r}

hypo_6 <- train[, .N , by = .(payment)]
hypo_6_b <- train[, .N, by = .(payment_type)]

hypo_6
hypo_6_b

```

We can see, that only the named payment or payment type differs in these two values. Therefore we can delete one of those values

```{r}

train[, payment_type := NULL]

```

## Hypothesis 7: The quanity is associated to the functionality of the pump.


```{r}

train[, quantity_group := NULL]

hypo_7 <- copy(train)

melted <- hypo_7[, .N, by = .(quantity, status_group)]
casted <- dcast(melted, quantity ~ status_group, value.var = "N")
casted <- casted %>%
  droplevels() %>%
  clean_names()


fig <- plot_ly(data = casted, x = ~quantity, y = ~functional, type = "bar", name = "functional") %>%
  add_trace(y = ~functional_needs_repair, name = "functional needs repair") %>%
  add_trace(y = ~non_functional, name = "non functional") %>%
  layout(yaxis = list(title = "Relative count"), barmode = "stack")

fig








```


We can see that dry pumps are almost exclusively non functional, while other quantities can differ between functional and non functional.




